---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "80%"
)
```

# simbased

<!-- badges: start -->
<!-- badges: end -->

`simbased` implements simulation-based inference as an alternative to the delta method for computing functions of model parameters, such as marginal effects. 

## Installation

You can install the development version of `simbased` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ngreifer/simbased")
```

## Example

Below is an example of performing g-computation for the ATT after logistic regression to compute the marginal risk ratio and its confidence interval:

```{r example, fig.width=7, fig.height=3}
library(simbased)

data("lalonde", package = "MatchIt")

fit <- glm(I(re78 == 0) ~ treat * (age + educ + race + married + nodegree + re74 + re75),
           data = lalonde, family = binomial)

set.seed(123)
sim_coefs <- sim(fit)

sim_est <- sim_apply(sim_coefs, function(fit) {
  d <- subset(lalonde, treat == 1)
  d$treat <- 1
  p1 <- mean(predict(fit, newdata = d, type = "response"))
  d$treat <- 0
  p0 <- mean(predict(fit, newdata = d, type = "response"))
  c(`E[Y(0)]` = p0, `E[Y(1)]` = p1, RR = p1 / p0)
}, verbose = FALSE)

summary(sim_est, null = c(NA, NA, 1))
sim_plot(sim_est)
```

We could have used `marginaleffects`, which uses the delta method instead:

```{r}
# Marginal risk ratio ATT, delta method-based
marginaleffects::comparisons(fit, variables = list(treat = 0:1),
                             newdata = subset(lalonde, treat == 1),
                             transform_pre = "lnratioavg") |>
  summary(transform_avg = exp)
```

The simulation-based confidence intervals are a bit wider, reflecting that the delta method may be a poor approximation.

`simbased` provides a shortcut for computing average marginal effects and comparisons between average adjusted predictions, `sim_ame()`, which is essentially a wrapper for `sim_apply()` with extra processing. We can compute the marginal risk ratio below:

```{r}
# Marginal risk ratio ATT, simulation-based
sim_est <- sim_ame(sim_coefs, var = "treat", subset = treat == 1,
                   contrast = "RR", verbose = FALSE)

summary(sim_est, null = c(NA, NA, 1))
```

We can also use `simbased` to compute predictions and first differences at set and typical values of the predictors, mimicking the functionality of `Zelig`'s `setx()` and `setx1()` functions, using `sim_setx()`:

```{r, fig.width=7, fig.height=3}
# Predictions across age and treat at typical values
# of the other predictors
sim_est <- sim_setx(sim_coefs, x = list(age = 20:50, treat = 0:1),
                    verbose = FALSE)

setx_plot(sim_est)
```

`simbased` offers parallel processing for all estimation functions to speed up computation.

The methodology of simulation-based inference is described in King, Tomz, and Wittenberg (2000).

King, G., Tomz, M., & Wittenberg, J. (2000). Making the Most of Statistical Analyses: Improving Interpretation and Presentation. *American Journal of Political Science*, 44(2), 347–361. https://doi.org/10.2307/2669316
